A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-03-02 paper/paper1/ms.tex
paper section on 1D parker winds

>>>>

\section{1D Parker wind models}
The dramatic leading tail of atmospheric escape suggests an underlying 6-dimensionality (3 space dimensions and 3 velocity dimensions), coupled with stellar wind \citep{2022ApJ...926..226M} may be needed to depict the problem faithfully.  MHD simulations are needed.  Here we explore one-dimensional (1D) models, which offer ease of interpretation and rapid calculation.  We employ the open-source \texttt{p-winds} code \citep{2022A&A...659A..62D}, which implements a transonic 1D Parker wind model with radiative transfer of the Helium 10830 triplet \citep{2018ApJ...855L..11O,2020A&A...636A..13L}.

The predicted Helium ionization depends sensitively on the XUV flux \citep{2019ApJ...881..133O}.  Few XUV measurements exist for F stars, with the F7IV-V \object[tau Boo]{$\tau$ Boo} serving as the closest analog \citep{2011A&A...532A...6S}.  \hatp resides closer to the Kraft break than does \object[tau Boo]{$\tau$ Boo}, which may be associated with rapid changes in the stellar wind and other atmospheric changes leading to much lower XUV flux.  We explore incident XUV fluxes of $100\times$ higher and lower than \object[tau Boo]{$\tau$ Boo} to gauge the sensitivity.  Figure XX shows the range of conceivable XUV SEDs.

Figure XX shows the 1D wind velocity and density computed with \texttt{p-winds}, assuming the baseline planet properties, a characteristic exosphere temperature of $T\sim10^4$ K, and including tidal gravity.
